---
title: "Budweiser"
author: "Bernard Strzalkowski and Keith Sikes"
date: '2022-06-18'
output: html_document
---

```{r map graphics}
if (!require("usmap", quietly = TRUE)) install.packages("usmap")
install.packages("BiocManager")

library(usmap)
library(stringr)
library(ggplot2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if("e1071" %in% rownames(installed.packages()) == FALSE) {
  install.packages("e1071")}
if("caTools" %in% rownames(installed.packages()) == FALSE) {
  install.packages("caTools")}
if("caret" %in% rownames(installed.packages()) == FALSE) {
  install.packages("caret")}
  if("moments" %in% rownames(installed.packages()) == FALSE) {
  install.packages("moments")}
    
library(readr)
library(dplyr)
library(tidyr)
require(e1071)
require(caTools)
require(caret)
library(class)

```

## Basics 
### Input, group_by, sum, change column names, merge

1.	How many breweries are present in each state?
2.	Merge beer data with the breweries data. Print the first 6how bservations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)

To make this code usable by anyone we pulled our data from github. We renamed the columns to merge the data and then grouped them by state.

```{r}
Breweries <- read_csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv")
Beers <- read_csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Beers.csv", col_types = cols(Beer_ID = col_integer(), 
    Brewery_id = col_integer()), na = "NA")

# rename columns
rename(Beers, Brew_ID = Brewery_id,Beer=Name) -> Beers
rename(Breweries, state = State,Brewery=Name) -> Breweries

# group by state, count breweries in state, order by count and state.
Q1 <- Breweries %>% group_by(state) %>% summarise(Density = n()) %>% arrange(desc(Density),state)

# merge many to one 
Q2 <- inner_join(Beers,Breweries,by ="Brew_ID")

head(Q2)
tail(Q2)

write_csv(Q1,file="StateBrew.csv")
```


## Choropleth

Here you see a map of the united states that shows the density of breweries through color code.

```{r}

plot_usmap(regions ="state", data = Q1, size=0.2, values = "Density", labels = TRUE) + 
  scale_fill_continuous( low = "white", high = "#0072B2", 
                         name = "Breweries", label = scales::comma) +
  theme(legend.position = "right") +
  labs(title = "Brewery Density by State") +
  theme(legend.position = "right") + 
  theme(legend.key.size = unit(2, 'cm'), legend.text=element_text(size=20)) + # adjust size of legend 
  theme(plot.title = element_text(hjust = 0.5))

```




### Median ABV

3.	Address the missing values in each column.
4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

In order to get rid of the N/A variables in the data we got the median value of Alcohol By Volume, or ABV, for each state and filled in the N/A variables with this value.

```{r}
medianABV = median(Q2$ABV)
medianABV

medianABVna = median(Q2$ABV, na.rm = TRUE)
medianABVna

Q2 %>% group_by(state) %>% 
  summarize(Median_ABV = median(ABV,na.rm=TRUE),
            Max_ABV = max(ABV, na.rm=TRUE)) -> StateABV
StateABV

# sorting
Q2 %>% arrange(desc(ABV))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match maxstateABV)
d <- ggplot(StateABV, aes(y=Median_ABV,x=reorder(as.factor(state),Median_ABV), fill = Median_ABV))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Median Alcohol By Volume Content By State") +
  xlab("State") +
  ylab("ABV") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))+
  theme(plot.title = element_text(hjust = 0.5))

d

```

### Median IBU

```{r}
medianIBU = median(Q2$IBU)
medianIBU

medianIBUna = median(Q2$IBU, na.rm = TRUE)
medianIBUna

Q2 %>% group_by(state) %>% 
  summarize(Median_IBU = median(IBU,na.rm=TRUE),
            Max_IBU = max(IBU, na.rm=TRUE)) -> StateIBU
StateIBU

# sorting
Q2 %>% arrange(desc(IBU))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match Median_ABV)
d <- ggplot(StateIBU, aes(y=Median_IBU,x=reorder(as.factor(state), Median_IBU), fill = Median_IBU))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Median International Bitterness Unit By State") +
  xlab("State") +
  ylab("IBU") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))+
  theme(plot.title = element_text(hjust = 0.5))

d

```

### Max ABV

5.	Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

Here we sorted the data in Q2 and created a separate data frame for the max ABV and IBU. We then made a bar chart displaying the results for both ABV and IBU. We also identified the max ABV and IBU across all "51" states.

```{r}
medianABV = median(Q2$ABV)
medianABV

medianABVna = median(Q2$ABV, na.rm = TRUE)
medianABVna

Q2 %>% group_by(state) %>% 
  summarize(Median_ABV = median(ABV,na.rm=TRUE),
            Max_ABV = max(ABV, na.rm=TRUE)) -> StateABV
StateABV

# sorting
Q2 %>% arrange(desc(ABV))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match maxstateABV)
d <- ggplot(StateABV, aes(y=Max_ABV,x=reorder(as.factor(state),Max_ABV), fill = Max_ABV))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Max Alcohol By Volume Content By State") +
  xlab("State") +
  ylab("ABV") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))+
  theme(plot.title = element_text(hjust = 0.5))

d

```
```{r}
medianIBU = median(Q2$IBU)
medianIBU

medianIBUna = median(Q2$IBU, na.rm = TRUE)
medianIBUna

Q2 %>% group_by(state) %>% 
  summarize(Median_IBU = median(IBU,na.rm=TRUE),
            Max_IBU = max(IBU, na.rm=TRUE)) -> StateIBU
StateABV

# sorting
Q2 %>% arrange(desc(IBU))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match maxstateABV)
d <- ggplot(StateIBU, aes(y=Max_IBU,x=reorder(as.factor(state),Max_IBU), fill = Max_IBU))
d <- d + geom_col(stat="identity") + coord_flip() +
  ggtitle("Max International Bitterness Unit By State") +
  xlab("State") +
  ylab("IBU") +
  theme(axis.text = element_text(colour = "black", size = 6)) +
  theme(panel.background = element_rect(fill = "white"))+
  theme(plot.title = element_text(hjust = 0.5))

d

```

#### Comparing ABV Data

6.	Comment on the summary statistics and distribution of the ABV variable.

Here we created a scatter plot for the max and median data for both ABV and IBU. We wanted to make sure the data was consistent so we reversed the max and median and created separate plots to identify trends.

```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

# Calculate Moving Average for Median
StateABV$MA5 <- TTR::SMA(StateABV$Median_ABV, n=5)

ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=Median_ABV)) +
  geom_smooth(method = "lm", colour = "darkblue") +
  ggtitle("Comparing Median to Max ABV") +
  xlab("Max ABV") +
  ylab("Median ABV") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))+
  theme(plot.title = element_text(hjust = 0.5))



```


```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

# Calculate Moving Average for Median
StateABV$MA5 <- TTR::SMA(StateABV$Median_ABV, n=5)

ggplot(data = StateABV, aes(x = Median_ABV, y = Max_ABV)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=Max_ABV)) +
  geom_smooth(method = "lm", colour = "darkblue") +
  ggtitle("Comparing Max to Median ABV") +
  xlab("Median ABV") +
  ylab("Max ABV") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))+
  theme(plot.title = element_text(hjust = 0.5))



```

```{r}
summary(Q2$ABV)
hist(Q2$ABV)

ggplot(data = Q2, aes(x = ABV)) + 
  geom_histogram(color="white", fill="#0072B2") +
  ggtitle("Summary of ABV") +
  xlab("ABV") +
  ylab("Frequency") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75)) +
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}

hist(Q2$ABV,
main="Range of Alcohol by Volume",
xlab="Alcohol by Volume",
xlim=c(0.02,0.10),
col="lightblue",
freq=TRUE
)

```

```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)
library(tidyverse)
library(tidyr)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

BeerComparison <- Q2
BeerComparison <- BeerComparison[!is.na(BeerComparison$IBU),]

ggplot(data = BeerComparison, aes(x = ABV, y = IBU)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=IBU)) +
  geom_smooth(method = "lm", colour = "darkblue") +
  ggtitle("IBU to ABV Beer Comparison") +
  xlab("ABV") +
  ylab("IBU") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))+
  theme(plot.title = element_text(hjust = 0.5))



```

```{r}
library(quantmod)
library(ggplot2)
library(TTR)
library(ggrepel)
library(tidyverse)
library(tidyr)

#plot(x = StateABV$Max_ABV, y = StateABV$Median_ABV, col = as.factor(StateABV$state))

# df <- ggplot(data = StateABV, aes(x = Max_ABV, y = Median_ABV, fill = as.factor(state)) + geom_point()

BeerComparison <- Q2
BeerComparison <- BeerComparison[!is.na(BeerComparison$IBU),]

ggplot(data = BeerComparison, aes(x = IBU, y = ABV)) + 
  scale_color_gradient(low = "cyan", high = "blue4") +
  geom_point(aes(colour=ABV)) +
  geom_smooth(method = "lm", colour = "darkblue") +
  ggtitle("ABV to IBU Beer Comparison") +
  xlab("IBU") +
  ylab("ABV") +
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))+
  theme(plot.title = element_text(hjust = 0.5))


```
7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

We created separate scatter plots to identify a relationship between bitterness and alcohol content. There seems to be a direct relationship between the two. This means that as one increases so does the other. The outliers are more than likely due to flavorings to make the beers sweeter or more bitter.

```{r}
model = lm(Q2$IBU~Q2$ABV)
plot(Q2$ABV, Q2$IBU,  main = "Are Alcohol Volume and Bitterness related?")
abline(coef(model))
list <- summary(model)
list
```

```{r}
list = grepl("ale",Q2$Style, ignore.case=TRUE)
alelist <- Q2 %>% filter(grepl("ale",Style, ignore.case=TRUE) & 
                           !(grepl("IPA",Style)|grepl("IPA",Beer)))
IPAlist <- Q2 %>% filter((grepl("IPA",Style)|grepl("IPA",Beer)))
alelist$type <- "ale"
IPAlist$type <- "IPA"
Beerlist <- rbind(alelist,IPAlist)
KNND <- as.factor(Beerlist$type)
KNNI <- Beerlist[,c(3,4)]
```

```{r}
BeerListClean <- Temp
BeerListClean <- na.omit(Beerlist)
KNNdataframe <- as.factor(BeerListClean$type)
KNNdataframe <- cbind(Type=KNNdataframe,BeerListClean[,c(3,4)])

# plot(KNNdataframe$IBU,KNNdataframe$ABV, col = KNNdataframe$Type, main = "Relationship of alcohol to bitterness in Ales and IPAs", ylab= "Alcohol percentage", xlab="International Bitterness Units")
# legend("bottomright", legend = paste("Type", 1:2), col = 1:3, pch = 19, bty = "n")

ggplot(KNNdataframe, aes(x=ABV,y=IBU, color=Type)) +
  scale_color_manual(values=c("cyan","blue4")) +
  geom_smooth(method = "lm") +
  geom_point() +
  labs(title = "Relationship of Alcohol and Bitterness in Ales and IPAs", y= "Alcohol percentage", x="International Bitterness Units") +
  theme(legend.position = "right") +   
  theme(axis.text = element_text(colour = "black", size = 10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid = element_line(size = 0.5, linetype = "solid", colour = "black")) +
  theme(panel.border = element_rect(colour = "black", fill=NA, size = 0.75))+
  theme(plot.title = element_text(hjust = 0.5))
```

###  KNN analysis 

8.	Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN classification to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand conceptually. 
In addition, while you have decided to use KNN to investigate this relationship (KNN is required) you may also feel free to supplement your response to this question with any other methods or techniques you have learned.  Creativity and alternative solutions are always encouraged.  

We used Knn analysis to predict whether or not a beer is an IPA or an Ale based on the ABV and IBU for each beer and we achieved between 80~90% accuracy with K=29

```{r multi knn, warning=FALSE}

# Creatubg the training and test datasets (70/30 split)
seed =(100)
set.seed(seed)
split <- sample.split(KNNdataframe, SplitRatio = 0.7)
train <- subset(KNNdataframe, split == "TRUE")
test <- subset(KNNdataframe, split == "FALSE")

# Feature Scaling and dataset splitting
trainknn <- scale(train[, 2:3])
testknn <- scale(test[, 2:3])# for knn model, need numeric
traincl <- train$Type
testcl <- test$Type

# do knn test - Using k=29 since it had the highest accuracy when running k=15 to k=30
knnhat <- knn(train=trainknn, test=testknn, cl=traincl, k=29)
knnCM <- table(testcl, knnhat)

#Show how accurate, sensitive, and specific the model is
CM = confusionMatrix(knnCM)
Acc =  CM$overall[1]
Sen =  CM$byClass[1]
Spec = CM$byClass[2]

cat(" Accuracy Sensitivity Specificity\n",Acc,Sen,Spec)
CM
```

### Extra = check out most popular style of beer by state

We separated out factors by style and state to find out how many were in each state. The most quantity then became the most popular for that state. We then used a color map diagram to demonstrate our findings on beer popularity

```{r Favs by state}
Favs = Q2[,c(2,6,10)]
Favs$Style =as.factor(Favs$Style)
result <- Favs %>% group_by(state,Style) %>% 
  summarise(Cnt = n()) %>% arrange(state,desc(Cnt)) %>%
  group_by(state) %>% filter(row_number()==1)
print(result)
```

``` {r socks off map}
p <- plot_usmap(regions ="state", data = result[,c(1,2)], size=.2, values = "Style", labels = TRUE) +   # plots map without state names
  scale_fill_brewer(palette = "RdYlBu") +
  labs(title = "Beer Preference by State") +
  theme(legend.position = "right") + 
  theme(legend.key.size = unit(.9, 'cm'), legend.text=element_text(size=12)) + # adjust size of legend 
  theme(plot.title = element_text(hjust = 0.5))

print(p)
```

